---
title: "Assignment 2 - SQL Connection"
author: "Dhairav Chhatbar"
date: "9/3/2019"
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Choose six recent popular movies.  Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5.  Take the results (observations) and store them in a SQL database.  Load the information into an R dataframe.

Your deliverables should include your SQL scripts and your R Markdown code, posted to GitHub.


## SQL Connection


```{r SQL Connection}
library(RMySQL)

drv2 <- dbDriver("MySQL")
con2 <- dbConnect(drv2, username="dsc", password="datascience", dbname ="movies", host="localhost")

dbListTables(con2)
```

## Querying Database
```{r Query}

dbGetQuery(con2, 'select Movies.Title, Reviews.UName as UserName, Reviews.Rating from
movies Join Reviews
Where Movies.VID = Reviews.VID')



```


## Enter into Data Frame

```{r Dataframe}
MovieReviews <- dbGetQuery(con2, 'select Movies.Title, Reviews.UName as UserName, Reviews.Rating from
movies Join Reviews
Where Movies.VID = Reviews.VID')

str(MovieReviews)

head(MovieReviews)

```
