---
title: "Assignment 3 - Regex"
author: "Dhairav Chhatbar"
date: "9/13/2019"
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 3

Copy the introductory example. The vector namestores the extracted names.

R> name

[1] "Moe Szyslak"       "Burns, C. Montgomery" "Rev. Timothy Lovejoy"

[4] "Ned Flanders"      "Simpson, Homer"       "Dr. Julius Hibbert"

**(a)  Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.**

```{r 3a}
library(stringr)
names <- c("Moe Szyslak", "Burns, C. Montgomery", "Rev. Timothy Lovejoy", "Ned Flanders", "Simpson, Homer", "Dr. Julius Hibbert")

names

first_name <- unlist(str_extract_all(names,"[[:alpha:]]+\\s|[.,] [[:alpha:]]+$"))
first_name <- unlist(str_extract_all(first_name, "[[:alpha:]]+"))
first_name

last_name <- str_extract(names, "([[:alpha:]]+?[,]|[[:alpha:]]+$)")
last_name <- str_extract(last_name, "[^,]+")
last_name

firstlast <- str_c(first_name, " ", last_name)
firstlast





```

**(b) Construct a logical vector indicating whether a character has a title (i.e.,Rev. and Dr.).**
```{r 3b}
title <-str_detect(names, "^[[:alpha:]]+?['.']")
title 
```

**(c) Construct a logical vector indicating whether a character has a second name.**
```{r 3c}
second_name <- str_detect(names, "[[:upper:]]['.']")
second_name
```

### Question 4
Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

**(a) [0-9]+\\$**
```{r 4a}
#Returns a string of numbrs of n length ending with the character $

s4a <- c("56424895314$", "534565645656", "dsdf542$65201" , "a442367", "b542346y", "abc123$0$ qw567$$")
unlist(str_extract_all(s4a, "[0-9]+\\$"))


```

**(b) \\b[a-z]{1,4}\\b**
```{r 4b}
#Returns a string of lowercase letters of length 1, 2, 3, or 4

s4b <- c("ahdb32ja", "a", "ab", "abc", "aBc", "abcd", "123abcd", "ab89@0", "371$12")
unlist(str_extract_all(s4b, "\\b[a-z]{1,4}\\b"))

```

**(c) .*?\\.txt$ **
```{r 4c}
#Returns a string of length n of any characters (except new line) ending in the text '.txt' or just the text '.txt'

s4c <- c("filename.txt", "file name.txt", "284@$.!.txt", "284@$.!.", "2file name.txt", ".txt", "abc", "2.txt")
unlist(str_extract_all(s4c, ".*?\\.txt$"))

```

**(d) \\d{2}/\\d{2}/\\d{4}**
```{r 4d}
#Returns a date of format xx/xx/xxxx where x is a while number
s4d <- c("42", "33/35/99", "01/18/2199", "33\22\1111")

unlist(str_extract_all(s4d, "\\d{2}/\\d{2}/\\d{4}"))

```

**(e) <(.+?)>.+?</\\1>**
```{r 4e}
#Returns a string what starts with the format <s1> s2 </s1>, where s1 is a string of any length and characters, and s2 is a string of any length. This string usually represents internet tags

s5e <- c("abcd", "4sdf asdf", "2", "  ", "abc xyz /abc", "<html> Hello World 1 </html>", "<a> b </c>", "<a> b </a>", "<b>  </b>", "<abc 123> b@2 </abc 123>")

str_extract_all(s5e, "<(.+?)>.+?</\\1>")

```

### Question 9
The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials at www.r-datacollection.com.

clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr

```{r 9}

unknowntext <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaotj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

#Remove all lower cases
knowntext <- unlist(str_extract_all(unknowntext, "[[:upper:]]|[[:punct:]]"))
knowntext

#combine into one string
knowntext <- str_c(knowntext, collapse = "")
knowntext

#Split string on the "." characters and put each word in individual element
knowntext <- unlist(str_split(knowntext, "\\."))
knowntext


```