---
title: "Assignment 1 - Mushrooms"
author: "Dhairav Chhatbar"
date: "8/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment

Very often, we’re tasked with taking data in one form and transforming it for easier downstream analysis. We will
spend several weeks in this course on tidying and transformation operations. Some of this work could be done in
SQL or R (or Python or…). Here, you are asked to use R—you may use base functions or packages as you like.
Mushrooms Dataset. A famous—if slightly moldy—dataset about mushrooms can be found in the UCI repository
here: <https://archive.ics.uci.edu/ml/datasets/Mushroom>. 

The fact that this is such a well-known dataset in the
data science community makes it a good dataset to use for comparative benchmarking. For example, if someone
was working to build a better decision tree algorithm (or other predictive classifier) to analyze categorical data,
this dataset could be useful. A typical problem (which is beyond the scope of this assignment!) is to answer the
question, “Which other attribute or attributes are the best predictors of whether a particular mushroom is
poisonous or edible?”

Your task is to study the dataset and the associated description of the data (i.e. “data dictionary”). You may need
to look around a bit, but it’s there! You should take the data, and create a data frame with a subset of the columns
in the dataset. You should include the column that indicates edible or poisonous and three or four other columns.
You should also add meaningful column names and replace the abbreviations used in the data—for example, in the
appropriate column, “e” might become “edible.” Your deliverable is the R code to perform these transformation
tasks.

## Import Data Dictionary & Dataset

Import mushrooms description and dataset from GitHub

```{r Import}
path <- 'https://raw.githubusercontent.com/dhairavc/DATA607/master/Assignment%201%20agaricus-lepiota.data'
path2 <- 'https://raw.githubusercontent.com/dhairavc/DATA607/master/Assignment1%20-%20agaricus-lepiota.names'

readLines(path2)

mushrooms <- read.csv(path)
head(mushrooms)
```

## Update Column Names
Give more meaningful header names to the dataset

```{r Names}
names(mushrooms) <- c("Class", "Cap-Shape", "Cap-Surface", "Cap-Color", "Bruises", "Odor", "Gill-Attachment", "Gill-Spacing", "Gill-Size", "Gill-Color", "Stalk-Shape","Stalk-Root","Stalk-Surface-Above-Ring","Stalk-Surface-Below-Ring","Stalk-Color-Above-Ring","Stalk-Color-Below-Ring","Veil-Type","Veil-Color","Ring-Number","Ring-Type","Spore-Print-Color","Population","Habitat")

head(mushrooms)


```

## Subset Dataframe
Create a subset of the original dataset

```{r Subset}

subshrooms <- mushrooms[c("Class", "Odor", "Population", "Habitat")]
head(subshrooms)

```

## Name Observations
Give useful names to observation values

``` {r Namesx}
str(subshrooms)
levels(subshrooms[,1]) <- c("edible", "poisonus")
levels(subshrooms[,2]) <- c("almond","creosote", "foul", "anise", "musty", "none", "pungent", "spicy", "fishy")
levels(subshrooms[,3]) <- c("abundant", "clustered", "numerous", "scattered", "several", "solitary")
levels(subshrooms[,4]) <- c("woods", "grasses", "leaves", "meadows", "paths", "urban", "waste")
str(subshrooms)


head(subshrooms)
```
