---
title: "Project 1"
author: "Dhairav Chhatbar"
date: "9/15/2019"
output:
  html_document: default
  pdf_document: default
---


### Project 1

In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:

Player’s Name

Player’s State

Total Number of Points

Player’s Pre-Rating

Average Pre Chess Rating of Opponents

For the first player, the information would be:

Gary Hua, ON, 6.0, 1794, 1605

1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.

### Import Data
Load text file into GitHub repository and read into R. Trim any trailing white spaces
```{r Import}
library(stringr)
path <- 'https://raw.githubusercontent.com/dhairavc/DATA607/master/tournamentinfo.txt'

chess_results <- readLines(path, warn = FALSE)
chess_results <- str_trim(chess_results, side = 'right')
chess_results

```

### Get Lines
Seperate data into into two patterns for extraction
``` {r Names}
Line1 <- grep(chess_results, pattern = "[[:space:]]{3,}[[:digit:]]{1,2}", value = TRUE)
Line1

Line2 <- grep(chess_results, pattern = "[[:space:]]{3}[[:alpha:]]{2}", value = TRUE)
Line2
```

### Get Player ID
Extract the Player ID. This ID can be used a primary identifier in the data for the player table
``` {r Player ID}
player_ID <- str_extract(Line2, "[[:digit:]]{8}")
chess_results_table <- data.frame(player_ID)
head(chess_results_table)

```

### Get Pair
Extract the player's pair number
``` {r pair}

pair_num <- str_extract(Line1, "^(.{5})")
pair_num <- str_trim(pair_num, side = 'both')

chess_results_table$Pair_Num <- as.numeric(pair_num)
#chess_results_table <- data.frame(as.numeric(pair_num))
#chess_results_table)[2] <- "Pair_Num"
head(chess_results_table)

```


### Get Names
Extract player names
``` {r P Names}
player_names <- str_extract(Line1, "\\|[[:space:]]([[:alpha:]]*[[[:space:]]|[[:punct:]]])*[[:space:]]")
player_names <- str_sub(player_names, 3, length(player_names))
player_names <- str_trim(player_names, side = 'right')
chess_results_table$Player_Name <- player_names
head(chess_results_table)

```

### Get State
Extract player state
```{r getState}
player_state <- str_extract(Line2, "[[:alpha:]]{2}")
chess_results_table$State <- player_state
head(chess_results_table)

```




### Get Points
Extract player points
``` {r points}
player_points <- str_extract(Line1, "[[:digit:]]\\.[[:digit:]]")
chess_results_table$Points <-  as.numeric(player_points)
head(chess_results_table)


```

### Get Pre_Rating
Extract each player's rating
``` {r pre Rating}
pre_rating <- str_extract(Line2, "R:[[:space:]]{1,2}[[:digit:]]{3,4}")
pre_rating <- str_extract(pre_rating, "[[:digit:]]{3,4}")
chess_results_table$Pre_Rating  <- as.numeric(pre_rating)
head(chess_results_table)
```



### Average Pre Chess Rating of Opponents
Based on the amount of matches each player has played, compute the average rating for all of each's player's opponents
``` {r Average Pre Chess Rating of Opponents, fig.width=52}
opps_played <- str_extract_all(Line1, "[W,L,D,U][[:space:]]{2}[[[:digit:]][[:space:]]][[:digit:]]")
opps <- suppressWarnings(str_extract_all(opps_played, "[[:digit:]]{1,2}"))
avg_rating <- vector()

for(i in 1:nrow(chess_results_table))
{
  avg_rating[i] <- mean(chess_results_table$Pre_Rating[as.integer(opps[[i]])])
}

chess_results_table$Opp_Avg_Rating <- avg_rating

```

### Final Table
Display the final table in a format that can be imported into a database
``` {r Finaltable, width = 8}
library(kableExtra)
kable(chess_results_table) %>% kable_styling(bootstrap_options = c("striped", "hover"))


```

### Write CSV
Produce a CSV file, which can be imporated into a database
```{r CSV}
write.csv(chess_results_table, file = "ChessResults.csv", row.names = FALSE )

```


